RSA Project Report Paper

### 2016-11-29 Paper/Paper.tex
Profread

	Quadratic Sieve method is actually built upon previous factorization method. It is a optimization of Dixon's Factorization Method. So, to learn what is Quadratic Sieve, we have to understand Dixon's Method first, which is also related to Fermat's Factorization and Kraitchik's Factorization Method. Here, we will briefly introduce these three factorization methods.

@@

	For example,

@@

	\renewcommand\labelenumi{\Roman{enumi}}
	\begin{enumerate}
		\item \textit{Choosing B \& M:} \itemgap
		\\ We read a lot of resources and most of them suggested a value between $\exp{(\frac{1}{2}\sqrt{\log{n}\log{\log{n}}})}$ and $\exp{(2\sqrt{\log{n}\log{\log{n}}})}$ (the reasoning behind choosing it this way is given in an explanation appendix A of Contini's paper \cite{contini}. At first in class, Professor Boklan emphasized that the number is just a toy number, therefore, we decided to use the smaller bound $\exp{(\frac{1}{2}\sqrt{\log{n}\log{\log{n}}})}$ instead. We then use this bound to test our number, resulting in 1.4 million and the a factor base of size about 50 thousand. However, since we used Pollard Rho to find an initial small factor, 809, we reduce the original number by dividing this factor out. We then recalculate again, resulting in 970000+, and a smaller factor base, shrinking to 40 thousand. For the convenience of communication, we choose 1 million as our B.

@@

		To decide whether $\legendre{N}{p} = 1$, we implemented both Euler criterion and Legendre Symbol. And probably because our factor base was not enough, we did not notice Legendre Symbol is much faster than Euler Criterion. Since professor suggest us use Legendre Symbol, we are using Legendre symbol to test $\legendre{N}{p}$.

		\item \textit{Pre-Sieving:} \itemgap
		\\ In this stage, we did a lot optimizations. First, we use an array as dictionary, which means we use prime valuse as the indices to the array, meaning since our B is one million, our array length is also one million. The reason why we do this is because we to be extremely fast in our sieving process. Since the array get operation is $\mathcal{O}(1)$ this is very fast. The downside, however, is obviously wasting memory. We then use this approach to save 3 informations. First one, is the roots of quadratic residue. Since every odd prime has two solutions, we use a two dimensional array to save both. Secondly, we also cache the value of $\log p$, since we do not want any unnecessary calculations in our sieving process. Finally, we also store the offset of x from the start position $\lfloor \sqrt{N} \rfloor - M$. We did not see any resource mention this, but we think it is useful, since we can very easily obtain the smooth numbers by just adding the offset to the starting position.

@@

		\item \textit{Sieving:} \itemgap
		\\ The sieving array is to store the accumulated $\log{p}$ value for each position in the range of [$-M, M$]. Therefore, its length is $2M+1$. The sieving process is just go through each prime in our factor base. Get its position which already cached in previous stage, and add $r+ip$ where $r$ is the position to the sieve array.

		This stage is probably most simple part of our program. It is not the idea is simple, it is made to be simple, since this is the bottle neck of our whole program. if we made this process calculates more, the program will slow down dramatically. And we also did this part in parallel. Not only multi-thread but also multi-process.

		We implemented a sever which only responsible to generate job which specify the range to search and add it to a task queue. And then, each client connected to sever and got job from the queue. The job could be very large, so we split the job according to the memory of client, and starting multiple threads according to the number of processors of client. Then, each thread search different ranges within the job.

		\item \textit{Save Sieving Result:} \itemgap
		\\ If we found a relation $\left|sieve\_array[position]-(0.5\log{N}+\log{M})\right| < threshold$ during the sieving, we consider this position is our smooth candidate, and save this position to our candidates set. And the reason we use the data structure of set is that it is possible to have duplicate positions and we want avoid checking process. We know for a set to check inclusion is very fast, since it just check whether the hash of target is in the table, which is almost $\mathcal{O}(1)$.

		When finish sieving, we go through each position we saved, and do trial division. We  only record the full relations, and the partial relations with prime leftover. And we use millar rabin testing check the primality of the leftover.

		And we also record the odd power prime factors, and save these informations to a concurrent set. Since this process will do in parallel, we have to keep in mind that concurrency would be a issue, therefore, we need to keep mutual exclusive when we do the save action.

		After collecting all the data, the client will sent it back to sever. And sever just simply save them in a file.

		\item \textit{Build Exponent Matrix:} \itemgap
		\\ We retrieve the file from sever, and parse it line by line. Here we met a problem that still confused us that is even though we collect data in different range , we find duplicates relations. This problem force us to collect more relations than we need. It could be a bug of our code. So, we need check duplication first.
		After that since each line in file correspond to a relation, we need to differentiate full or partial relations. if it is full relation we just save $x$ in a list and all the prime factors of $Q(x)$ in a dictionary, prime factor as key, and exponent as value.(why we save this will be explained in \textit{Calculating Factors}). Since we have the odd exponent factors information of the relation, it is very simple to use it building a exponent vector and save it in matrix. if it is partial relation, we save it a dictionary first, and use the leftover as a key. until we find another partial relation have the same leftover, we pop out it from the dictionary, and combine two relation together as following. First, we multiply $x_1$ and $x_2$ as $x$, then, merge the two prime factor dictionaries. Finally, we xor two exponent vectors and save it in matrix.

		Since we did not use the meataxe, we need introduce our matrix implementation. The reason why we were not using meataxe is that it is not very well documented, and sometimes it does not work as we expected. We are afraid of spending too many time on it. And another reason is our number is not insanely large, it just 40000x40000 matrix.

		And the reason why it is hard to manipulate large matrix is our ram not big enough to save it. Let us do a simple calculation, 40000x40000=1600000000 elements. If we use a integer to represent each element that will cost 1600000000*4=6400000000 bytes $\approx$ 6Gb ram. This is too memory consuming. However, since the matrix is over $\mathcal{F}_2$, we can use binary representation for each element it will only cost 6Gb/32 $\approx$ 190Mb which is acceptable.

@@

		First, we make a very simple header, which only has two integers that specific the row number and the column number. Then, we use 1 byte to represent 8 bits, and if the matrix's column size mod 8 is not 0, we are padding zeros after each row. For example 01 becomes 01000000 which is 40 in hex. This is for sake of simplicity of implementation. And finally, we just save each row linearly.

		To access certain rows, say the 8th row, we have our initial position pointer $s$, and we calculate many bytes in each row by row\_length = (column+padding)/8, and the start position of 8th row is 8*row\_length+$s$.

		To access certain column, say the 6th column, we calculate this column in which byte by 6/8=0, this means 6th column belongs to the first byte, then, take the byte and right shift 6 position and do a ``and'' operation with 1 which means ignore other bits except for the last bit.

@@

		We use Gaussian elimination to find the null space. The reason why we can use Gaussian elimination to find null space, is because any elementary row operation can be seen as a matrix transformation. For example.

@@

		So, if we reduce $M$ to its reduced echelon form $M^\prime$ by Gaussian Elimination, means that $\exists E=\prod_{i=1}^k{E_i}\ s.t.\ EM \rightarrow M^\prime$, where $E_i$ is the elementary row operation. We also know at the bottom of reduced echelon form always have all zero rows, and that is the null space.

		Our approach is to records all the elementary operation by using a identity matrix $I$ since $EI=E$.

@@

		First, we will going to explain why finding null space is equal to find the sum of rows equal to zero. It is easy only if you notice that and none of the resources we read explained this. It is just because the matrix production is for each row multiply a coefficient add rows together. And we are working over $\mathbb{F}_2$, the multiplier 0 means not using this row, 1 means using this row. And this is the reason why we are checking each element of the vector in null space, if we found a 1, we find the corresponding row in matrix and get related $x$ and all the prime factors of $Q(x)$.

		Recall that in \textit{Building Exponent Matrix} Section, we said we will explain why we save all the prime factors of $Q(x)$ instead of just saving $Q(x)$ itself. The reason is that if we multiply these $Q(x)$ , we get an insanely large number, we do not know any good algorithm to find the square root of it, we tried newton's square root approximation, which did not help us much. We can not get a accurate number. Therefore, we saved all the prime factor and their exponents. And then, we only need to half of the exponent and multiply them together we get the $\sqrt{\prod Q(x)}$.

		After we got $\sqrt{\prod Q(x)}$, our final answer is $gcd\left(\prod Q(x) \pm \prod x, N\right)$, if we are lucky not all relation we've found are trivial, we have one factor of N.

@@

	\section {Collecting data by parallel computing} 
	\subsection{Computer spec}
	computer 1: i7-4700M, 2.4GHz (boost 3.1Ghz), RAM 16G\\
	\indent computer 2: i7-3537U, 2.0GHz               , RAM 8G\\
	\indent computer 3: i5-3570K, 3.4GHz               , RAM 16G\\
	\indent computer 4: i5-4210U, 1.7GHz               , RAM 8G

	\subsection{Data collection}
	We collect 706854 relations in 3 hours and 20 minutes with 4 parallel computing computers. In the 706854 relations, there are 48059 full relations, 305914 partial relations and the remaining are duplicate relations.
	Thus, there are about 85$\%$ of the relations are partial relations. In other words, when modulus N grows bigger and bigger, the probabilities of finding full relations gets lower and lower.

@@

	Once we got the factors of N, we use Euler-totient function calculates the $\varphi(N)$, and then we can calculate the private key $d$ by find the mod inverse of $e \mod{\varphi(N)}$ thorough extend euclidean gcd algorithm. After we got $d$ we can recover the message just by taking $(m^e)^d \pmod{n}$.

	Then, we got a message encoded as 1121225152171520201541591920182520182511292020125\\2051445181451919. Since, we know it is English and the encoding method already, we use three methods to recover it to English very quickly.

	\renewcommand\labelenumi{\arabic{enumi}.}
	\begin{enumerate}
		\item We use a English dictionary, and encodes the words exactly the same way as the message did. And we only searching for the longer words(at least 4 letters). And we extract the matching part, which could shrink the length of the message. In our case, we use this method found the word ``tenderness''.

		\item We wrote a program to calculate all the possible combinations. The idea is using backtrack algorithm to build a tree. Then, saving all the leaves in the tree. Finally, collecting each path from the leaf to root. The example for code 21525 in below:

@@

		\item Since the message is not very long and encoding method is easy, we just use our English knowledge to test whether the words make sense in its place.

@@

	And finally, we solve this assignment and decrypt message M shown as following:\\

>>>>

	Quadratic Sieve method is actually built upon a previous factorization method. It is a optimization of Dixon's Factorization Method. So, to learn what is Quadratic Sieve, we have to understand Dixon's Method first, which is also related to Fermat's Factorization and Kraitchik's Factorization Method. Here, we will briefly introduce these three factorization methods.

@@

@@

	\renewcommand\labelenumi{\Roman{enumi}}
	\begin{enumerate}
		\item \textit{Choosing B \& M:} \itemgap
		\\ We read a lot of resources and most of them suggested a value between $\exp{(\frac{1}{2}\sqrt{\log{n}\log{\log{n}}})}$ and $\exp{(2\sqrt{\log{n}\log{\log{n}}})}$ (the reasoning behind choosing it this way is given in an explanation appendix A of Contini's paper \cite{contini}. At first in class, Professor Boklan emphasized that the number is just a toy number, therefore, we decided to use the smaller bound $\exp{(\frac{1}{2}\sqrt{\log{n}\log{\log{n}}})}$ instead. We then use this bound to test our number, resulting in B being 1.4 million and the a factor base of size about 50 thousand. However, since we used Pollard Rho to find an initial small factor, 809, we reduce the original number by dividing this factor out. We then recalculate again, resulting in B of 970000+, and a smaller factor base, shrinking to 40 thousand. For the convenience of communication, we choose 1 million as our B.

@@

		Again, to decide whether $\legendre{N}{p} = 1$, we implemented both Euler's Criterion and Legendre Symbol and again probably because our factor base was not that large, we did not notice that much of a difference in the Law of Quadratic Reciprocity over Euler's Criterion. Since Professor Boklan, suggested us to use Law of Quadratic Reciprocity, we ended up using this to test $\legendre{N}{p}$.

		\item \textit{Pre-Sieving:} \itemgap
		\\ In this stage, we did a lot optimizations. First, we use an array as a dictionary, which means we use prime valuse as the indices to the array, meaning since our B is one million, our array length is also one million. The reason why we do this is because we to be extremely fast in our sieving process. Since the array get operation is $\mathcal{O}(1)$ this is very fast. The downside, however, is obviously wasting memory. We then use this approach to save 3 informations. First one, is the roots of quadratic residue. Since every odd prime has two solutions, we use a two dimensional array to save both. Secondly, we also cache the value of $\log p$, since we do not want any unnecessary calculations in our sieving process. Finally, we also store the offset of x from the start position $\lfloor \sqrt{N} \rfloor - M$. We did not see any resource mention this, but we think it is useful, since we can very easily obtain the smooth numbers by just adding the offset to the starting position.

@@

		\item \textit{Sieving:} \itemgap
		\\ The sieving array is used to store the accumulated $\log{p}$ value for each position in the range of [$-M, M$]. Therefore, its length is $2M+1$. The sieving process goes as follows:\\(1) For each prime in our factor base, $p$, get its position which should already be cached in the pre-sieving stage\\(2) Add $r+ip$ where $r$ is the position in the sieve array for $i\ge0$ but $r+ip<M$.

		This stage is probably the most simple part of our program. However, it is not the idea that is simple, but it was made to be simple, since this is the bottle neck of our whole program. If we made this process calculate more, the program will slow down dramatically. Additionally, we did this part in parallel, utilizing multi-threading but also multi-processing.

		We implemented a server whose sole responsible was to generate jobs. Each job specifies the range in which we search for smooth candidates. Each job is then added to a task queue in which each client who is connected to server will take a job off the queue. Additionally, the jobs can be very large, so we split the jobs according to the memory of the client; starting multiple threads according to the number of processors the client has in which each thread searches a different range within the job.

		\item \textit{Save Sieving Result:} \itemgap
		\\ If we found a relation $\left|sieve\_array[position]-(0.5\log{N}+\log{M})\right| < threshold$ during the sieving, we consider this position a smooth candidate, and save this position to our candidates set. The reason we use the set data structure is that it is possible to have duplicate positions and we want to avoid a checking process to remove those duplicates. We know that for sets, to check inclusion is very fast, since it is just to check whether the hash of target is in the table, which is almost $\mathcal{O}(1)$.

		When finish sieving, we go through each position we saved, and do trial division. We only record the full relations, and the partial relations with primes leftover. In order to check for the primality of the leftover primes, we use Millar-Rabin primality testing.

		In addition, we need to record the odd power prime factors, and save this information to a concurrent set. Since this process will be done in parallel, we have to keep in mind that concurrency would be a issue, therefore, we need to keep mutual exclusiveness when we do the saving action.

		After collecting all the data, the client will send it back to server, which the server then simply saves in a file.

		\item \textit{Build Exponent Matrix:} \itemgap
		\\ We retrieve the file from server, and parse it line by line. Here we faced a problem that still confuses us; even though we collect data in different ranges, we still found duplicate relations, though the reason behind these duplicates could just be a bug in the code. This problem forced us to collect more relations than we need which we then had to be wary of duplicates when parsing.  
		Afterwards, since each line in the file corresponds to a relation, we need to differentiate full or partial relations. If it is full relation we just save $x$ in a list and all the prime factors of $Q(x)$ in a dictionary: using a prime factor as key, and exponent as value.(why we saved them this will be explained in \textit{Calculating Factors}). Since we have information on the odd exponent factors of the relation, it is very simple to use it building a exponent vector and save it in the matrix. If it is partial relation, we save it to a dictionary first, and use the leftover as a key. Until we find another partial relation(having the same leftover), we pop it from the dictionary, and combine the two relations together as follows. First, we multiply $x_1$ and $x_2$ as $x$, then, merge the two prime factor dictionaries. Finally, we xor the two exponent vectors and save it in the matrix.

		Since we did not use the meataxe, we need to introduce our matrix implementation. The reason why we did not using meataxe is that it is not very well documented, and sometimes it does not work as we expected. We were afraid of spending too much time on it, and seeing that our 40000x40000 matrix is not insanely large.

		The reason why it is hard to manipulate large matrices is because our RAM is not large enough to save it. Let us do a simple calculation, 40000x40000=1600000000 elements. If we use a integer to represent each element that will cost 1600000000*4=6400000000 bytes $\approx$ 6Gb ram. This is too large of a memory consumption. However, notice since the matrix is over $\mathcal{F}_2$, we can use binary representation for each element it will only cost 6Gb/32 $\approx$ 190Mb which is acceptable.

@@

		First, we make a very simple header, which only has two integers that specific the number of rows and columns. Then, we use 1 byte to represent 8 bits, and if the matrix's column size mod 8 is not 0, we are padding zeros after each row. For example 01 becomes 01000000 which is 40 in hex. This is for sake of simplicity of implementation. Finally, we just save each row linearly.

		To access certain rows, say the 8th row, we have our initial position pointer $s$, and we calculate how many bytes in each row, row\_length = (column+padding)/8, in which we can find the start position of 8th row, 8*row\_length+$s$.

		To access certain columns, say the 6th column, we calculate in a similar way. A column in which byte by 6/8=0, this means 6th column belongs to the first byte, then, take the byte and right shift 6 position and do a ``and'' operation with 1 which means to ignore other bits except for the last bit.

@@

		We use Gaussian elimination to find the null space. The reason why we can use Gaussian elimination to find null space, is because any elementary row operation can be seen as a matrix transformation. For example:

@@

		So, if we reduce $M$ to its reduced echelon form $M^\prime$ by Gaussian Elimination, which means $\exists E=\prod_{i=1}^k{E_i}\ s.t.\ EM \rightarrow M^\prime$, where $E_i$ is the elementary row operation. We also know that at the bottom of the reduced echelon form matrix, it will always have all zero rows, which is the null space.

		Our approach is to record all the elementary operations by using a identity matrix $I$ since $EI=E$.

@@

		First, we will explain why finding the null space is equal to finding the sum of rows equal to zero, because initially we did not realize this ourselves from all the resources we read. Recall the how we perform the product of the matrices: for each row multiply a coefficient add rows together. We are working over $\mathbb{F}_2$, the multiplier 0 means not using this row, 1 means using this row, which is why we are checking each element of the vector in the null space; if we found a 1, we find the corresponding row in matrix and get related $x$ and all the prime factors of $Q(x)$.

		Recall that in \textit{Building Exponent Matrix} Section, we said we will explain why we save all the prime factors of $Q(x)$ instead of just saving $Q(x)$ itself. The reason is that if we multiply those $Q(x)$ , we get an insanely large number, which we do not know of any good algorithm to find the square root of. We tried Newton's Square Root approximation, which did not help us much, seeing as we were unable to get an accurate enough of a number. Therefore, we saved all the prime factors and their exponents, which we then only have to halve the exponents and multiply them together,getting $\sqrt{\prod Q(x)}$.

		After we got $\sqrt{\prod Q(x)}$, our final answer is $gcd\left(\prod Q(x) \pm \prod x, N\right)$. If we are lucky and not all of the relations we find are trivial, we have finally the factors of N.

@@

	\section {Collecting data by parallel computing} 
	\subsection{Computer spec}
	Computer 1: i7-4700M, 2.4GHz (overclocked to 3.1Ghz), RAM 16G\\
	\indent Computer 2: i7-3537U, 2.0GHz               , RAM 8G\\
	\indent Computer 3: i5-3570K, 3.4GHz               , RAM 16G\\
	\indent Computer 4: i5-4210U, 1.7GHz               , RAM 8G

	\subsection{Data collection}
	We collect 706854 relations in 3 hours and 20 minutes with 4 computers computing in parallel. In the 706854 relations, there are 48059 unique full relations, and 305914 unique partial relations. The remaining 352,881 relations are duplicate relations.
	Additionally, approximately 85$\%$ of the relations are partial relations. In other words, when our x grows larger and larger, the probability of finding full relations gets lower and lower.

@@

	Once we got the factors of N, calculate the $\varphi(N)$, and then we can calculate the private key $d$ by find the mod inverse of $e \mod{\varphi(N)}$ through the Extend Euclidean Algorithm. After we get $d$ we can recover the message just by taking $(m^e)^d \pmod{n}$.

	Ultimately, we get a message encoded as 1121225152171520201541591920182520182511292020125\\2051445181451919. Since, we know it is English and the encoding method already, we use three methods to recover it to English very quickly.

	\renewcommand\labelenumi{\arabic{enumi}.}
	\begin{enumerate}
		\item We use an English dictionary, and encode the words exactly the same way as the message did. We then only search for the longer words(at least 4 letters) and we extract the matching part, which could shrink the length of the message. In our case, using this method, we found the word ``tenderness''.

		\item We wrote a program to calculate all the possible combinations. The idea is using backtrack algorithm to build a tree ,then saving all the leaves in the tree. Finally, we collect each path from the a leaf to the root. The example for code 21525 in below:

@@

		\item Since the message is not very long and encoding method is easy, we just use some human intuition and English knowledge to test whether the words make sense in its place.

@@

	Finally, with much effort, we were finally able to retrieve and decrypt the original message, M, shown as following:\\
